\mytitle

### 2012-03-15 papers/SAS2012/implicitization.tex
on ne distingue plus X^s et X^d, et l'algo de calcul des nouveaux chemins a changé

Guided static analysis, as proposed by \cite{DBLP:conf/sas/GopanR07}, applies to the transition
graph of the program. We now present a new technique applying this analysis on the implicit
multigraph from \cite{Monniaux_Gonnord_SAS11}, thus avoiding control flow merges with
unfeasible paths.
In this section, we use the same notations as \S\ref{sec:path_focusing}, except
that we call the abstract values $X_i^s$ instead of $X_i$ to leave room for
another abstract value $X_i^d$ later.

@@

We distinguish three phases in the main loop of the analysis:
\begin{compactenum}
\item \label{step:addingpaths} We start finding a new relevant subset
  $P \cup P'$ of the graph.
  Either the previous iteration or the initialization led us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  $P'$. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} in Algorithm~\ref{algo:combined}.
\item \label{step:ascending} Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  Algorithm~\ref{algo:combined}.
\item  \label{step:narrowing} We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more points to explore,
  i.e. $A' = \emptyset$.
\end{compactenum}

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}^s$ by the path transformation
is not included in the current $X_{j}^s$.
Intuitively, such a path makes the abstract value $X_{j}^s$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i^s$ the abstract value
associated to $p_i$ :
\vspace{-.5em}
\[
f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
 \neg X_j^s)
\vspace{-.1em}
\]
The difference with~\cite{Monniaux_Gonnord_SAS11} is that we do not
work on the entire transition graph but on a subset of it. Therefore we
conjoin the formula $f(p_i)$ with the actual set of working paths,
noted $P$, expressed as a Boolean formula, where the Boolean variables are the
\emph{reachability predicates} of the control points. We can easily construct
this formula from the binary decision diagram using dynamic programming, and
avoiding an exponentially sized formula. In other words, we force the SMT solver
to give us a path included in $P$.

@@

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.
The paths in $P'$ are computed one after another, until no more path
can make the invariant grow. This is line~\ref{alg=computeNewPaths} in
Algorithm~\ref{algo:combined}, which corresponds to
Algorithm~\ref{algo:computepaths}. We also use SMT solving to discover
these new paths, but we subtly change the SMT formula given to the
SMT solver: we now simply check for the satisfiability of $f(p_i)$
instead of $f(p_i) \wedge P$. Since we already know that $f(p_i)
\wedge P$ is unsatisfiable, none of the paths given by the SMT solver
will be in $P$.

% À changer si on décide de ne plus distinguer $X^s$ et $X^d$
Also, to prevent the algorithm from adding too many paths at a time,
and in particular to prevent the algorithm from trying to re-add the
same path infinitely many times, we use another abstract value
associated to the control point $p_j$, noted $X_j^d$, which is distinct
from $X_j^s$, and initialized to $X_j^s$ right before computing new paths. 
In the SMT formula, we associate to $p_j$ the abstract value $X_j^s$ when $p_j$
is the starting point of the path, and $X_j^d$ when it is its destination point.
We thus check the satisfiability of the formula
$f'(p_i)$, where:
\vspace{-.5em}
\[
f'(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge X_i^s \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
\neg X_j^d)
\vspace{-1em}
\]
$X_j^d$ is updated when the point $p_j$ is the target of a new path.
This way, further SMT queries do not
compute other paths with the same source and destination if it is not
needed (because these new paths would not make $X_j^d$ grow, hence
would not be returned by the SMT solver).

@@

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{inparaenum}
\item 
the subset of paths $P$ strictly increases at each loop iteration,
and is bounded by the finite set of
paths in the entire graph. 
\item the set $P'$ always verifies $P \cap P' = \emptyset$ by construction, 
which guarantees that $P'$
will eventually be empty after a finite number of loop iterations.
\end{inparaenum}

>>>>

Guided static analysis, as proposed by \cite{DBLP:conf/sas/GopanR07}, applies to the transition
graph of the program. We now present a new technique applying this analysis on the implicit
multigraph from \cite{Monniaux_Gonnord_SAS11}, thus avoiding control flow merges with
unfeasible paths.
In this section, we use the same notations as \S\ref{sec:path_focusing}.
%, except
%that we call the abstract values $X_i^s$ instead of $X_i$ to leave room for
%another abstract value $X_i^d$ later.

@@

We distinguish three phases in the main loop of the analysis:
\begin{compactenum}
\item \label{step:addingpaths} We start finding a new relevant subset
  $P$ of the graph.
  Either the previous iteration or the initialization led us to a
  state where there are no more paths in the previous subset $P$,
  starting at $p_i$, that make the abstract values of the successors
  grow (otherwise, the SMT solver would not have answered
  ``\emph{unsat}''). Narrowing iterations preserve this property.
  However, there may exist such paths in the entire multigraph, that
  are not in $P$. This phase computes these paths and adds them to
  the subset. This phase is described in~\ref{subsec:addingpaths}
  and corresponds to lines in~\ref{alg=start-add-paths} to
  \ref{alg=end-add-paths} in Algorithm~\ref{algo:combined}.
\item \label{step:ascending} Given a new subset $P$, we search for paths starting at point
  $p_i \in P_R$, such that these paths are in $P$, i.e are included in
  the working subgraph. Each time we find a path, we update the
  abstract value of the destination point of the path. This is the
  phase explained in~\ref{subsec:ascending}, and corresponds to
  lines~\ref{alg=start-ascending} to~\ref{alg=end-ascending} in
  Algorithm~\ref{algo:combined}.
\item  \label{step:narrowing} We perform narrowing iterations the usual way
  (line~\ref{alg=narrowing} in algorithm~\ref{algo:combined}) and
  reiterate from step 1 unless there are no more points to explore,
  i.e. $A' = \emptyset$.
\end{compactenum}

@@

In order to find which path to focus on, we construct an SMT formula $f(p_i)$, whose
model when satisfiable is a path that starts in $p_i$, goes to a successor $p_j
\in P_R$ of $p_i$, such that the image of $X_{i}$ by the path transformation
is not included in the current $X_{j}$.
Intuitively, such a path makes the abstract value $X_{j}$ grow, and thus is
an interesting path to focus on. We loop until the formula becomes unsatisfiable,
meaning that the analysis of $p_i$ is finished.

If we note $Succ(i)$ the set of indices $j$ such that $p_j \in P_R$ is a
successor of $p_i$ in the expanded multigraph, and $X_i$ the abstract value
associated to $p_i$ :
\vspace{-.5em}
\[
f(p_i) = \rho \wedge b_i^s \wedge 
\displaystyle\bigwedge_{j \in P_R \atop j \neq i} \neg
b_j^s \wedge X_i \wedge \displaystyle\bigvee_{j \in Succ(i)} (b_j^d \wedge
 \neg X_j)
\vspace{-.1em}
\]
The difference with~\cite{Monniaux_Gonnord_SAS11} is that we do not
work on the entire transition graph but on a subset of it. Therefore we
conjoin the formula $f(p_i)$ with the actual set of working paths,
noted $P$, expressed as a Boolean formula, where the Boolean variables are the
\emph{reachability predicates} of the control points. We can easily construct
this formula from the binary decision diagram using dynamic programming, and
avoiding an exponentially sized formula. In other words, we force the SMT solver
to give us a path included in $P$.
Each time the invariant candidate of a point $p_j$ has been updated, $p_j$ is
inserted into $A'$ since it may be the start of a new feasible paths.

@@

This new subgraph is the union of the paths of the previous one with a set
$P'$ of new paths that become feasible regarding the current abstract values.
The paths in $P'$ are computed one after another, until no more path
can make the invariant grow. This is line~\ref{alg=computeNewPaths} in
Algorithm~\ref{algo:combined}, which corresponds to
Algorithm~\ref{algo:computepaths}. We also use SMT solving to discover
these new paths, but we subtly change the SMT formula given to the
SMT solver: we now try to find a path that is not yet in $P$, but is feasible
and makes the invariant candidate of its destination grow.
% À changer si on décide de ne plus distinguer $X^s$ et $X^d$
We thus check the satisfiability of the formula
$f'(p_i)$, where:
\vspace{-.5em}
\[
f'(p_i) = f(p_i)\wedge \neg P
\vspace{-1em}
\]
$X_j$ is updated using an abstract union when the point $p_j$ is the target of a new path. This way, further SMT queries do not
compute other paths with the same source and destination if it is not needed
(because these new paths would not make $X_j$ grow, hence would not be returned
by the SMT solver).

@@

When a new path has been found, it is immediately added into $P$. 
We then have to add $p_i$ and $p_j$ into $A$ (since we do not apply widening in
this section) and $p_j$ into $A'$, since $p_j$ may be the starting point of a
new feasible path.

\subsection{Termination}
Termination of this algorithm is guaranteed, because:
\begin{inparaenum}
\item 
the subset of paths $P$ strictly increases at each loop iteration,
and is bounded by the finite set of
paths in the entire graph. 
\item
	When computing new paths, we cunjunct our formula with $\neg P$, meaning
	that we obtain each possible path only once. The number of path is finite,
	so this computation always terminates.
\item
	The Path-focusing iterations terminate because of the properties of
	widening.
\end{inparaenum}
